Analysis-ready VCF at Biobank scale using Zarr

### 2025-01-22 paper.tex
Add missing data and update table format

Therefore, to reduce storage
requirements we generated a Zarr schema for encoding that uses local
allele fields with \texttt{vcf2zarr mkschema --local-alleles}, and
used the generated schema as input to \texttt{vcf2zarr encode}. Zarr conversion required 14h09min runtime, generating a dataset with 35 arrays, consuming a total of 2.3TiB storage over 11,235,556 chunk files, with a reduction in storage of 3.2X. Therefore, by using local allele fields we achieve approximately 3X improvement of compression. The top fields in terms of storage are detailed in Table~\ref{tab-spruce-data}.

\begin{table}
\caption{Summary for a selection of the largest VCF Zarr columns produced for 
1,063 VCFs on 12 spruce chromosomes using \texttt{vcf2zarr} with samples chunk size set to 1,063 , variants chunk size to 10,000, and using local allele fields to reduce output size. For column details see the caption of Table~\ref{tab-genomics-england-data}.
\label{tab-spruce-data}}

\begin{tabular}{llS[table-format=3.1]S[table-format=3.2]S[table-format=3.2]}
\toprule
Field & type & storage & compress & percentage \\
\midrule
/call_LPL & int16 & 1.26 TiB & 17.0 & 54.76% \\
/call_LA & int8 & 415.39 GiB & 18.0 & 17.63% \\
/call_genotype & int8 & 381.28 GiB & 19.0 & 16.18% \\
/call_genotype_mask & bool & 103.84 GiB & 71.0 & 4.41% \\
/variant_DP4 & int32 & 26.24 GiB & 2.1 & 1.11% \\
/variant_MQSBZ & float32 & 13.08 GiB & 1.1 & 0.56% \\
/variant_RPBZ & float32 & 13.04 GiB & 1.1 & 0.55% \\
/variant_MQBZ & float32 & 12.97 GiB & 1.1 & 0.55% \\
/variant_SCBZ & float32 & 12.94 GiB & 1.1 & 0.55% \\
/variant_VDB & float32 & 12.84 GiB & 1.1 & 0.55% \\
/variant_quality & float32 & 12.84 GiB & 1.1 & 0.55% \\
/variant_BQBZ & float32 & 12.64 GiB & 1.1 & 0.54% \\
/variant_SGB & float32 & 12.12 GiB & 1.2 & 0.51% \\
/variant_position & int32 & 9.65 GiB & 1.4 & 0.41% \\
/variant_AC & int16 & 7.42 GiB & 2.8 & 0.31% \\
/variant_DP & int32 & 6.71 GiB & 2.1 & 0.28% \\
/variant_allele & object & 5.37 GiB & 21.0 & 0.23% \\
/variant_AN & int16 & 3.05 GiB & 2.3 & 0.13% \\
/call_genotype_phased & bool & 1.75 GiB & 2100.0 & 0.07% \\
/variant_filter & bool & 1.46 GiB & 2.4 & 0.06% \\
/variant_MQ & int8 & 489.41 MiB & 7.3 & 0.02% \\
/variant_IMF & float32 & 160.95 MiB & 89.0 & 0.01% \\
/variant_IDV & int16 & 120.72 MiB & 59.0 & 0.01% \\
\bottomrule
\end{tabular}
\end{table}

>>>>

Therefore, to reduce storage requirements we generated a Zarr schema
for encoding that uses local allele fields with \texttt{vcf2zarr
mkschema --local-alleles}, and used the generated schema as input to
\texttt{vcf2zarr encode}. Zarr conversion required 14h09min runtime,
generating a dataset with 35 arrays, consuming a total of 2.3TiB
storage over 11,235,556 chunk files, with a reduction in storage of
3.2X. Therefore, by using local allele fields we achieve approximately
3X improvement of compression. The top fields in terms of storage are
detailed inTable~\ref{tab-spruce-data}.

\begin{table}
\caption{Summary for a selection of the largest VCF Zarr columns
produced for 1,063 VCFs on 12 spruce chromosomes using
\texttt{vcf2zarr} with samples chunk size set to 1,063, variants
chunk size to 10,000, and using local allele fields to reduce output
size. For column details see the caption of
Table~\ref{tab-genomics-england-data}.
\label{tab-spruce-data}}

\begin{tabular}{llS[table-format=3.2]S[table-format=3.2]S[table-format=3.2]}
\toprule
{Field} & {type} & {storage} & {compress} & {\%total} \\
\midrule
/call\_LPL & int16 & 1.26 T & 17.0 & 54.76\% \\
/call\_LA & int8 & 415.39 G & 18.0 & 17.63\% \\
/call\_genotype & int8 & 381.28 G & 19.0 & 16.18\% \\
/call\_genotype\_mask & bool & 103.84 G & 71.0 & 4.41\% \\
/variant\_DP4 & int32 & 26.24 G & 2.1 & 1.11\% \\
/variant\_MQSBZ & float32 & 13.08 G & 1.1 & 0.56\% \\
/variant\_RPBZ & float32 & 13.04 G & 1.1 & 0.55\% \\
/variant\_MQBZ & float32 & 12.97 G & 1.1 & 0.55\% \\
/variant\_SCBZ & float32 & 12.94 G & 1.1 & 0.55\% \\
/variant\_VDB & float32 & 12.84 G & 1.1 & 0.55\% \\
/variant\_quality & float32 & 12.84 G & 1.1 & 0.55\% \\
/variant\_BQBZ & float32 & 12.64 G & 1.1 & 0.54\% \\
/variant\_SGB & float32 & 12.12 G & 1.2 & 0.51\% \\
/variant\_position & int32 & 9.65 G & 1.4 & 0.41\% \\
/variant\_AC & int16 & 7.42 G & 2.8 & 0.31\% \\
/variant\_DP & int32 & 6.71 G & 2.1 & 0.28\% \\
/variant\_allele & object & 5.37 G & 21.0 & 0.23\% \\
/variant\_AN & int16 & 3.05 G & 2.3 & 0.13\% \\
/call\_genotype\_phased & bool & 1.75 G & 2100.0 & 0.07\% \\
/variant\_filter & bool & 1.46 G & 2.4 & 0.06\% \\
/variant\_MQ & int8 & 489.41 M & 7.3 & 0.02\% \\
/variant\_IMF & float32 & 160.95 M & 89.0 & 0.01\% \\
/variant\_IDV & int16 & 120.72 M & 59.0 & 0.01\% \\
% /variant\_length & int16 & 105.19 M & 68.0 & 0.00\% \\
% /variant\_INDEL & bool & 72.18 M & 49.0 & 0.00\% \\
% /variant\_id & object & 34.53 M & 830.0 & 0.00\% \\
% /variant\_id_mask & bool & 32.74 M & 110.0 & 0.00\% \\
% /variant\_contig & int16 & 32.02 M & 220.0 & 0.00\% \\
% /variant\_FS & float32 & 31.67 M & 450.0 & 0.00\% \\
% /variant\_MQ0F & float32 & 31.67 M & 450.0 & 0.00\% \\
% /region\_index & int32 & 3.14 M & 2.7 & 0.00\% \\
% /contig\_length & int64 & 6.77 K & 1.2 & 0.00\% \\
% /contig\_id & object & 6.2 K & 1.3 & 0.00\% \\
% /sample\_id & object & 5.35 K & 1.6 & 0.00\% \\
% /filter\_id & object & 4.43 K & 0.0 & 0.00\% \\
\bottomrule
\end{tabular}
\end{table}
